### 2016-08-20 latex/paper.tex
Simplified K_rim expresion
Minor edits to sections 0-2

Here we derive such a formula. The non-dimensionalized buckling tension depends only on two parameters which are identified as the ratio of torsional stiffness to bending stiffness, and the ratio of spoke stiffness to bending stiffness. The number of ``waves'' in the buckled shape is determined by two stiffness ratios and may in general by greater than 2. Finally, we show that the wheel may be represented by a system of equivalent springs, two series-connected springs for the rim bending and torsion stiffness, and a parallel-connected spring for the spoke system.

\begin{keywords}
bicycle wheel, 
buckling, 
bifurcation,
rider safety.
\end{keywords}

@@

\section{DEFORMATION OF THE WHEEL}
A typical bicycle wheel (Fig. \ref{fig:schematic} consists of a slender rim connected to a hub by means of a system of very slender metal spokes. The spokes are usually threaded at the rim side into spoke nipples which are inserted through the rim from the outside. The spokes are kept under tension by the rim. Without pretension, the spokes would immediately buckle under a very small load due to their high aspect ratio.

@@

\begin{figure}[!ht]
\centering
\includegraphics[scale=1.0]{figures/bmd_figures-06.eps}
\caption{Deformation of a rim cross-section. The shear center $S\rightarrow S'$ translates by the vector $\bm{u} = u\e + v\ee + w\eee$ and rotates through an angle $\phi$.}
\label{fig:def}
\end{figure}

The in-plane curvature $\kappa_1$ is assumed to remain constant during flexural-torsional buckling. The in-plane curvature, out-of-plane curvature and twist rate are
	\begin{equation}\label{eq:k1}
	\kappa_1 = 1/R
	\end{equation}
	\begin{equation}\label{eq:k2}
	\kappa_2 = u'' - \frac{\phi}{R}
	\end{equation}
	\begin{equation}\label{eq:k3}
	\kappa_3 = \phi' + \frac{1}{R} u'
	\end{equation}
The symbol $()'$ indicates a derivative with respect to $s$. It can easily be verified that Equations \ref{eq:k2} and \ref{eq:k3} simplify to the standard equations for straight beams in the limit $R\rightarrow \infty$. The curvature and twist rate produce a bending moment and twisting moment equal to
	\begin{equation}\label{eq:M2}
	M_2 = EI_{11} \kappa_2
	\end{equation}
	\begin{equation}\label{eq:M3}
	M_3 = GJ \kappa_3 + EI_w \kappa_3''
	\end{equation}
The total strain energy is decomposed into lateral bending, uniform torsion, and warping terms which are related to the curvatures.
	\begin{equation}\label{eq:Urim}
	U_{rim} = \int_0^{2\pi} \left( \frac{1}{2} EI_{22} \kappa_2^2 + \frac{1}{2} GJ \kappa_3^2 + \frac{1}{2} EI_w (\kappa_3')^2 \right)\, ds
	\end{equation}
After buckling, a differential line segment $ds$ deforms to $dS$. The apparent shortening of the line segment along the $s$ direction is then
	\begin{equation}\label{eq:ds}
	ds - dS = ds - \sqrt{ds^2 - \left(\frac{du}{ds}ds\right)^2} \approx \frac{1}{2} (u')^2 \, ds
	\end{equation}
The axial compressive force in the rim moves through this differential displacement, performing virtual work equal to
	\begin{equation}\label{eq:Vrim}
	V_{rim} = \frac{1}{2} \int_0^{2\pi} N_r (u')^2 \, ds
	\end{equation}

@@

In this paper we consider bicycle wheels with slender prestressed spokes. We model a single spoke as an elastic bar pinned at each end. This ensures that the spoke force is always collinear with the spoke axis. We fix our global coordinate system to the hub, which is assumed to be rigid. Therefore, the displacement of the hub node of a spoke is zero. The linearized elongation of a spoke is given by
	\begin{equation}\label{eq:selong}
	\Delta l = \bm{u}_n\cdot \n
	\end{equation}
where $\bm{u}_n$ is the displacement vector of the spoke nipple and \n{} is a unit vector pointing from the spoke nipple to the hub eyelet. The linearized rotation of the spoke is given by
	\begin{equation}\label{eq:srot1}
	\Omega_1 = \frac{1}{l} \bm{u}_n\cdot \nn
	\end{equation}
	\begin{equation}\label{eq:srot2}
	\Omega_2 = \frac{1}{l} \bm{u}_n\cdot \nnn
	\end{equation}
where \nn{} and \nnn{} are two orthogonal unit vectors orthogonal to \n{}, and $l$ is the spoke length.

The elongation produces a net force on the rim parallel to the original spoke axis equal to
	\begin{equation}\label{eq:sF1}
	f_1 = E_sA_s\left(\frac{\Delta l}{l}\right)\n = \frac{E_sA_s}{l} \n \cdot \bm{u}_n 
	\end{equation}
The rotation of the spoke produces net forces on the rim perpendicular to the original spoke axis due to the rotation of the initial tension vector.
	\begin{equation}\label{eq:sF2}
	f_2 = T \sin{\Omega_1} \approx \frac{T}{l} \nn \cdot \bm{u}_n
	\end{equation}
	\begin{equation}\label{eq:sF3}
	f_3 = T \sin{\Omega_2} \approx \frac{T}{l} \nn \cdot \bm{u}_n
	\end{equation}
since \n{}, \nn{}, \nnn{} are mutually orthogonal, the total force on the rim is given by $\bm{f} = \bm{k}_f \cdot \bm{u}_n$, where the spoke stiffness matrix is given by
	\begin{equation}\label{eq:kf}
	\bm{k}_f = \frac{E_sA_s}{l}\n\n + \frac{T}{l}(\nn\nn + \nnn\nnn)
	\end{equation}
The first term in Eqn.~\ref{eq:kf} is the {\bf elastic stiffness} and the second term is the {\bf tension stiffness}. The tension stiffness is the component responsible for the transverse vibrations of a guitar string, e.g.

Following the approach of Smith~\cite{Smith1901a} and Pippard~\cite{Pippard1932d}, we will approximate the spoke system as a continuum which exerts a distributed load on the rim. The stiffness per unit length is defined by the sum of the spoke stiffnesses in one periodic grouping of $n_p$ spokes, divided by the arc length of the grouping.
	\begin{equation}\label{eq:kbar}
	\bar{\bm{k}} = \frac{1}{2\pi R}\left(\frac{n_s}{n_p}\right) \sum_i^{n_p} \bm{k}_{f, i}
	\end{equation}
	The distributed load $\bar{\bm{f}}$ does an amount of work per unit length equal to
	\begin{equation}\label{eq:w_spokes}
	\bar{w} = \int_0^{u_n} \bar{\bm{f}} \cdot d\bm{u}_n = \frac{1}{2} \bm{u}_n \cdot \bar{\bm{k}} \cdot \bm{u}_n
	\end{equation}
The line of action is assumed to pass sufficiently close to the rim shear center that any difference between the displacement of the spoke nipple and the displacement of the shear center can be ignored. Therefore $\bm{u}_n = \bm{u}$. Substituting a lateral buckling displacement $\bm{u}=u\e$ into Eqn. \ref{eq:w_spokes} and integrating over the length of the rim gives the strain energy stored in the spoke system.
	\begin{equation}\label{eq:Us}
	U_{spokes} = \int_0^{2\pi} \bar{w} \, ds = \frac{1}{2} \bar{k}_{uu}u^2
	\end{equation}
The continuum spoke stiffness can be written $\bar{\bm{k}} = \bar{\bm{k}}^{el} + T\bar{\bm{k}}^{tens}$, where the dependence on $T$ is shown explicitly. A very close approximation for the $\bar{k}_{uu}$ component for a symmetrically dished wheel is given by
	\begin{equation}\label{eq:kuu}
	\bar{k}_{uu} = \frac{n_sE_sA_s}{2\pi Rl}\sin^2{\alpha} + \frac{n_s T}{2\pi Rl}
	\end{equation}

@@

The buckled shape of the wheel must be periodic, so we assume a buckled shape of the form
	\begin{equation}\label{eq:modeshape}
	\begin{split}
	u &= \delta u_n \cos{\frac{ns}{R}} \\
	\phi &= \delta\phi_n \cos{\frac{ns}{R}}
	\end{split}
	\end{equation}
The wheel becomes unstable when the increase in strain energy is exactly balanced by the virtual work for a perturbing displacement $\delta u_n, \delta\phi_n$, i.e. when the total potential energy is zero.
	\begin{equation}\label{eq:TotPot}
	\Pi = U_{rim} + U_{spokes} - V_{rim} = 0
	\end{equation}
Substituting the buckling mode shape Eqn. \ref{eq:modeshape} into Eqns. \ref{eq:Urim}, \ref{eq:Vrim}, and \ref{eq:Us} and substituting into Eqn. \ref{eq:TotPot} yields a buckling criterion for the $n$th mode.
	\begin{multline}\label{eq:ECrit}
	\frac{\pi EI_{11}}{R^3}(n^2 u_n - R\phi_n)^2 + \frac{\pi n^2}{R^2}\left(GJ + \frac{EI_w}{R^2}n^2\right)(u_n-R\phi_n)^2 + \\
	\pi R(\bar{k}_{uu}^{el} + T\bar{k}_{uu}^{el})u_n^2 - N_r\frac{\pi n^2}{R}u_n^2=0
	\end{multline}
The torsional rigidity and warping rigidity can be combined into an effective torsional stiffness $\widetilde{GJ} = GJ + n^2EI_w/R^2$. Equation \ref{eq:ECrit} has a quadratic form and can be written using matrix notation:
	\begin{equation}\label{eq:Qform}
	\begin{bmatrix}
	u_n\\\phi_n
	\end{bmatrix} \cdot
	\begin{bmatrix}
	\frac{\partial^2 \Pi}{\partial u_n^2} & \frac{\partial^2 \Pi}{\partial u_n \partial \phi_n}\\
	\frac{\partial^2 \Pi}{\partial u_n\partial\phi_n} & \frac{\partial^2 \Pi}{\partial \phi_n^2}
	\end{bmatrix} \cdot
	\begin{bmatrix}
	u_n\\\phi_n
	\end{bmatrix}
	=0
	\end{equation}
Non-trivial solutions of Eqn. \ref{eq:Qform} exist when the determinant of the matrix vanishes. Expanding the determinant in Eqn. \ref{eq:Qform} and substituting Eqn. \ref{eq:TN} yields a linear equation for the critical buckling tension of the $n$th mode.
	\begin{multline}\label{eq:Tc_eq}
	\left(\frac{2\pi EI_{22}}{R}+\frac{2\pi \widetilde{GJ}n^2}{R} \right) \left( \frac{2\pi \widetilde{GJ}n^2}{R^3} + \frac{2\pi EI_{22}n^4}{R^3} + 2\pi R\bar{k}_{uu}^{el} + \frac{n_sT_{cn}}{l} - \frac{n_sn^2T_{cn}}{R}\right)\\
	- \left( \frac{2\pi EI_{22}n^2}{R^2} + \frac{2\pi \widetilde{GJ}n^2}{R^2} \right)^2 = 0
	\end{multline}
Solving Eqn. \ref{eq:Tc_eq} for $T_{cn}$ yields the tension at which the $n$th mode becomes unstable. The critical buckling tension $T_c$ is the minimum $T_{cn}$ with respect to the mode number $n\geq 2$.

@@

Defining an equivalent torsion stiffness, bending stiffness, and spoke stiffness, the mode stiffness $P/u_n$ can be obtained from Eqn. \ref{eq:un} as
	\begin{equation}\label{eq:Kn_spring}
	K_n = \frac{K_{bending}K_{torsion}}{K_{bending}+K_{torsion}} + K_{spokes}
	\end{equation}
\,\,\,\, where
	\begin{equation}\label{eq:Ktorsion}
	K_{torsion} = \widetilde{GJ}n^2\frac{\pi (n^2-1)^2}{R^3}
	\end{equation}
	\begin{equation}\label{eq:Kbending}
	K_{torsion} = EI_{11}\frac{\pi (n^2-1)^2}{R^3}
	\end{equation}
	\begin{equation}\label{eq:Kspokes}
	K_{spokes} = \pi R \bar{k}_{uu}
	\end{equation}
The form of Equation \ref{eq:Kn_spring} suggests an equivalent spring model (Fig. \ref{fig:Kn}) for the $n$th mode in which the rim stiffness is represented by two springs connected which is then connected in parallel with the spoke stiffness. If there is a large difference between the rim bending stiffness and torsion stiffness, the equivalent rim stiffness will be dominated by the smaller of the two.

With this definition, Eqn. \ref{eq:Tc_nondim} has an extremely simple form
	\begin{equation}\label{eq:Tc_Kn}
	T_{cn} = \frac{2RK_n}{n_s(n^2 - R/l)}
	\end{equation}
In this form, the benefit of adding spoke stiffness becomes readily apparent. Because the rim stiffness and spoke stiffness act in parallel, increasing either one directly impacts the maximum tension.

@@

\begin{tabular}{c|p{12cm}}
\hline
\textit{\textbf{Term}} & \textit{\textbf{Definition}} \\
\hline
$n_s$		& Number of spokes\\
$E_s, A_s$	& Spoke Young's modulus and cross-sectional area\\
$l$			& Spoke length\\
$\alpha$	& Out-of-plane spoke angle--wall angle of a hypothetical ``cone'' of spokes\\
$\beta$		& In-plane spoke angle--measured between the in-plane projection of the spoke and
			  a radius of the wheel\\
$\n$		& Unit vector aligned with the spoke axis\\
$\nn,\nnn$	& Two unit vectors perpendicular to the spoke axis\\
$\e,\ee,\eee$ & Coordinate axes defined at the rim shear center\\
$\bm{u}$	& Displacement of the rim shear center\\
$u,v,w$		& Vector components of $\bm{u}$ in the \e{}, \ee{}, \eee{} directions.\\
$\phi$		& Rotation of the rim cross-section about the $\hat{\bm{e}}_3$ axis.\\
$\delta u, \delta \phi$ & Virtual displacement and rotation\\
$u_n, \phi_n$ & Buckling mode coefficients\\
$n$			& Buckling mode number\\
$T$			& Tension in a spoke\\
$T_{cn}$	& Critical tension for the $n$th mode\\
$T_c$		& Critical buckling tension (minimum of $T_{cn}$ over all modes)\\
$T_e$		& ``Euler'' tension--tension which would produce an axial force in the rim
			  equal to the buckling load for a straight beam-column of length $2\pi R$\\
$\bm{f}$	& Force exerted by a spoke on the rim\\
$f_1, f_2,f_3$ & Vector components of $\bm{f}$\\
$\bm{k}_f$	& Stiffness matrix for a single spoke\\
$\bar{\bm{k}}$ & Equivalent stiffness per-unit-length of the spoke system. $\bar{\bm{k}} = \bar{\bm{k}}^{el} + \bar{\bm{k}}^{tens}$\\
$\bar{\bm{k}}^{el}$ & Part of $\bar{\bm{k}}$ proportional to $E_sA_s$\\
$\bar{\bm{k}}^{tens}$ & Part of $\bar{\bm{k}}$ proportional to $T$\\
$\bar{k}_{ij}$ & Tensor components of $\bar{\bm{k}}$\\
$R$			& Rim radius--measured from the rim shear center\\
$E$			& Rim Young's modulus\\
$I_{11}$	& Second moment of area of the rim in the lateral direction\\
$J$			& Rim torsion constant\\
$I_w$		& Rim warping constant\\
$\widetilde{GJ}$ & Effective torsional rigidity including both Saint-Venant torsion and warping\\
$\tilde{\mu}$ & Ratio of effective torsional rigidity to bending rigidity: $\widetilde{GJ}/EI_{11}$\\
$\lambda$	& Ratio of spoke system stiffness to rim bending stiffness OR ratio of rim radius to 
			  the characteristic length of a beam on an elastic foundation.\\
\hline
\end{tabular}

>>>>

Here we derive such a formula. The non-dimensionalized buckling tension depends only on two parameters which are identified as the ratio of torsional stiffness to bending stiffness, and the ratio of spoke stiffness to bending stiffness. The number of ``waves'' in the buckled shape is determined by these two stiffness ratios and may in general by greater than 2. Finally, we show that the wheel may be represented by a system of equivalent springs: two series-connected springs for the rim bending and torsion stiffness, connected in parallel with a spring for the spoke system.

\begin{keywords}
bicycle wheel, 
buckling, 
bifurcation,
bicycle safety.
\end{keywords}

@@

\section{DEFORMATION OF THE WHEEL}
A typical bicycle wheel (Fig. \ref{fig:schematic}) consists of a slender rim connected to a hub by means of a system of very slender metal spokes. The spokes are usually threaded at the rim side into spoke nipples which are inserted through the rim from the outside. The spokes are kept under tension by the rim. Without pretension, the spokes would immediately buckle under a very small load due to their high aspect ratio.

@@

\begin{figure}[!ht]
\centering
\includegraphics[scale=1.0]{figures/bmd_figures-06.eps}
\caption{Rim cross-section in the undeformed and deformed configuration. The shear center $S\rightarrow S'$ translates by the vector $\bm{u} = u\e + v\ee + w\eee$ and rotates through an angle $\phi$.}
\label{fig:def}
\end{figure}

The in-plane curvature $\kappa_1$ remains constant during flexural-torsional buckling. The in-plane curvature, out-of-plane curvature and twist rate are
	\begin{equation}\label{eq:k1}
	\kappa_1 = 1/R
	\end{equation}
	\begin{equation}\label{eq:k2}
	\kappa_2 = u'' - \frac{\phi}{R}
	\end{equation}
	\begin{equation}\label{eq:k3}
	\kappa_3 = \phi' + \frac{1}{R} u'
	\end{equation}
The symbol $()'$ indicates a derivative with respect to arc-length $s$. It can easily be verified that Equations \ref{eq:k2} and \ref{eq:k3} simplify to the standard equations for straight beams in the limit $R\rightarrow \infty$. The total strain energy is decomposed into lateral bending, uniform torsion, and warping terms which are related to the curvatures.
% The curvature and twist rate produce a bending moment and twisting moment equal to
	% \begin{equation}\label{eq:M2}
	% M_2 = EI_{11} \kappa_2
	% \end{equation}
	% \begin{equation}\label{eq:M3}
	% M_3 = GJ \kappa_3 + EI_w \kappa_3''
	% \end{equation}
	\begin{equation}\label{eq:Urim}
	U_{rim} = \frac{1}{2} \int_0^{2\pi} \left( EI_{11} \kappa_2^2 + GJ \kappa_3^2 + EI_w (\kappa_3')^2 \right)\, ds
	\end{equation}
The length of a line segment of the rim is assumed to remain constant during buckling. The apparent shortening of an initial line segment $ds$ is given by the difference between $ds$ and $ds'$, the in-plane component of the deformed line segment.
	\begin{equation}\label{eq:ds}
	ds - ds' = ds - \sqrt{ds^2 - \left(\frac{du}{ds}ds\right)^2} \approx \frac{1}{2} (u')^2 \, ds
	\end{equation}
The axial compressive force in the rim moves through this differential displacement, performing virtual work equal to
	\begin{equation}\label{eq:Vrim}
	V_{rim} = \frac{1}{2} \int_0^{2\pi} N_r (u')^2 \, ds
	\end{equation}

@@

In this paper we consider bicycle wheels with slender prestressed spokes. We model a single spoke as an elastic bar pinned at each end, ensuring that the spoke force is always collinear with the spoke axis. We fix our global coordinate system to the hub, which is assumed to be rigid. The linearized elongation of a spoke is given by
	\begin{equation}\label{eq:selong}
	\Delta l = \bm{u}_n\cdot \n
	\end{equation}
where $\bm{u}_n$ is the displacement vector of the spoke nipple and \n{} is a unit vector pointing from the spoke nipple to the hub eyelet. The linearized rotation of the spoke is given by
	\begin{equation}\label{eq:srot1}
	\Omega_1 = \frac{1}{l} \bm{u}_n\cdot \nn
	\end{equation}
	\begin{equation}\label{eq:srot2}
	\Omega_2 = \frac{1}{l} \bm{u}_n\cdot \nnn
	\end{equation}
where \nn{} and \nnn{} are two orthogonal unit vectors orthogonal to \n{}, and $l$ is the spoke length. The elongation produces a net force on the rim parallel to the original spoke axis equal to
	\begin{equation}\label{eq:sF1}
	f_1 = E_sA_s\left(\frac{\Delta l}{l}\right)\n = \frac{E_sA_s}{l} \n \cdot \bm{u}_n 
	\end{equation}
The rotation of the spoke produces net forces on the rim perpendicular to the original spoke axis due to the rotation of the initial tension vector.
	\begin{equation}\label{eq:sF2}
	f_2 = T \sin{\Omega_1} \approx \frac{T}{l} \nn \cdot \bm{u}_n
	\end{equation}
	\begin{equation}\label{eq:sF3}
	f_3 = T \sin{\Omega_2} \approx \frac{T}{l} \nn \cdot \bm{u}_n
	\end{equation}
since \n{}, \nn{}, \nnn{} are mutually orthogonal, the total force on the rim is given by $\bm{f} = \bm{k}_f \cdot \bm{u}_n$, where the spoke stiffness matrix is given by
	\begin{equation}\label{eq:kf}
	\bm{k}_f = \frac{E_sA_s}{l}\n\n + \frac{T}{l}(\nn\nn + \nnn\nnn)
	\end{equation}
The first term in Eqn.~\ref{eq:kf} is the {\bf elastic stiffness} and the second term is the {\bf tension stiffness}. The tension stiffness is the component responsible for the transverse vibrations of a guitar string, e.g. The term $\n\n$ is the dyadic product, or tensor product (also called the outer product), and is defined by the property $\n\n\cdot u = (\n\cdot u)\n$.

Formalizing the approach of Smith~\cite{Smith1901a} and Pippard~\cite{Pippard1932d}, we will approximate the spoke system as a continuum which exerts a distributed load on the rim. The stiffness per unit length is defined by the sum of the spoke stiffnesses in one periodic grouping of $n_p$ spokes, divided by the arc length of the grouping.
	\begin{equation}\label{eq:kbar}
	\bar{\bm{k}} = \frac{1}{2\pi R}\left(\frac{n_s}{n_p}\right) \sum_i^{n_p} \bm{k}_{f, i}
	\end{equation}
	The distributed load $\bar{\bm{f}}$ does an amount of work per unit length equal to
	\begin{equation}\label{eq:w_spokes}
	\bar{w} = \int_0^{\bm{u}_n} \bar{\bm{f}} \cdot d\bm{u}_n = \frac{1}{2} \bm{u}_n \cdot \bar{\bm{k}} \cdot \bm{u}_n
	\end{equation}
The line of action is assumed to pass sufficiently close to the rim shear center that any difference between the displacement of the spoke nipple and the displacement of the shear center can be ignored. Therefore $\bm{u}_n = \bm{u}$. Substituting a lateral buckling displacement $\bm{u}=u\e$ into Eqn. \ref{eq:w_spokes} and integrating over the length of the rim gives the strain energy stored in the spoke system.
	\begin{equation}\label{eq:Us}
	U_{spokes} = \int_0^{2\pi} \bar{w} \, ds = \frac{1}{2} \bar{k}_{uu}u^2
	\end{equation}
The continuum spoke stiffness can be written $\bar{\bm{k}} = \bar{\bm{k}}^{el} + T\bar{\bm{k}}^{tens}$, where the dependence on $T$ is shown explicitly. A very close approximation for the $\bar{k}_{uu}$ component for a symmetrically dished wheel is given by
	\begin{equation}\label{eq:kuu}
	\bar{k}_{uu} = \frac{n_sE_sA_s}{2\pi Rl}\sin^2{\alpha} + \frac{n_s T}{2\pi Rl}
	\end{equation}

@@

The buckled shape of the wheel must be periodic, so we assume a buckled shape of the form
	\begin{equation}\label{eq:modeshape}
	\begin{split}
	u &= \delta u_n \cos{\frac{ns}{R}} \\
	\phi &= \delta\phi_n \cos{\frac{ns}{R}}
	\end{split}
	\end{equation}
The wheel becomes unstable when the increase in strain energy is exactly balanced by the virtual work for a perturbing displacement $\delta u, \delta\phi$, i.e. when the total potential energy is zero.
	\begin{equation}\label{eq:TotPot}
	\Pi = U_{rim} + U_{spokes} - V_{rim} = 0
	\end{equation}
Substituting the buckling mode shape Eqn. \ref{eq:modeshape} into Eqns. \ref{eq:Urim}, \ref{eq:Vrim}, and \ref{eq:Us} and substituting into Eqn. \ref{eq:TotPot} yields a buckling criterion for the $n$th mode.
	\begin{multline}\label{eq:ECrit}
	\frac{\pi EI_{11}}{R^3}(n^2 \delta u_n - R\delta\phi_n)^2 + \frac{\pi n^2}{R^2}\left(GJ + \frac{EI_w}{R^2}n^2\right)(\delta u_n-R\delta\phi_n)^2 + \\
	\pi R(\bar{k}_{uu}^{el} + T\bar{k}_{uu}^{el})\delta u_n^2 - N_r\frac{\pi n^2}{R}\delta u_n^2=0
	\end{multline}
The torsional stiffness and warping stiffness can be combined into an effective torsional stiffness $\widetilde{GJ} = GJ + n^2EI_w/R^2$. Equation \ref{eq:ECrit} has a quadratic form and can be written in matrix notation:
	\begin{equation}\label{eq:Qform}
	\begin{bmatrix}
	\delta u_n\\\delta\phi_n
	\end{bmatrix} \cdot
	\begin{bmatrix}
	\frac{\partial^2 \Pi}{\partial u_n^2} & \frac{\partial^2 \Pi}{\partial u_n \partial \phi_n}\\
	\frac{\partial^2 \Pi}{\partial u_n\partial\phi_n} & \frac{\partial^2 \Pi}{\partial \phi_n^2}
	\end{bmatrix} \cdot
	\begin{bmatrix}
	\delta u_n\\\delta\phi_n
	\end{bmatrix}
	=0
	\end{equation}
Non-trivial solutions of Eqn. \ref{eq:Qform} exist when the determinant of the matrix vanishes. Expanding the determinant in Eqn. \ref{eq:Qform} and substituting Eqn. \ref{eq:TN} yields a linear equation for the critical buckling tension of the $n$th mode.
	\begin{multline}\label{eq:Tc_eq}
	\left(\frac{2\pi EI_{22}}{R}+\frac{2\pi \widetilde{GJ}n^2}{R} \right) \left( \frac{2\pi \widetilde{GJ}n^2}{R^3} + \frac{2\pi EI_{22}n^4}{R^3} + 2\pi R\bar{k}_{uu}^{el} + \frac{n_sT_{cn}}{l} - \frac{n_sn^2T_{cn}}{R}\right)\\
	- \left( \frac{2\pi EI_{22}n^2}{R^2} + \frac{2\pi \widetilde{GJ}n^2}{R^2} \right)^2 = 0
	\end{multline}
Solving Eqn. \ref{eq:Tc_eq} for $T_{cn}$ yields the tension at which the $n$th mode becomes unstable. The critical buckling tension $T_c$ is the minimum $T_{cn}$ with respect to the mode number $n\geq 2$.

@@

Defining an equivalent torsion stiffness, bending stiffness, and spoke stiffness, the mode stiffness $P/u_n$ can be obtained from Eqn. \ref{eq:un} as
	\begin{equation}\label{eq:Kn_spring}
	K_n = \frac{K_{bending}K_{torsion}}{K_{bending}+K_{torsion}} + K_{spokes}
	\end{equation}
\,\,\,\, where
	\begin{equation}\label{eq:Ktorsion}
	K_{torsion} = \frac{\pi\widetilde{GJ}n^2}{R^3}(n^2-1)^2
	\end{equation}
	\begin{equation}\label{eq:Kbending}
	K_{torsion} = \frac{\pi EI_{11}}{R^3}(n^2-1)^2
	\end{equation}
	\begin{equation}\label{eq:Kspokes}
	K_{spokes} = \pi R \bar{k}_{uu}
	\end{equation}
The form of Equation \ref{eq:Kn_spring} suggests an equivalent spring model (Fig. \ref{fig:Kn}) for the $n$th mode in which the rim stiffness is represented by two springs connected which is then connected in parallel with the spoke stiffness. If there is a large difference between the rim bending stiffness and torsion stiffness, the equivalent rim stiffness will be dominated by the smaller of the two.

With this definition, the critical buckling tension for the $n$th mode (Eqn. \ref{eq:Tc_nondim}) has an extremely simple form
	\begin{equation}\label{eq:Tc_Kn}
	T_{cn} = \frac{2RK_n}{n_s(n^2 - R/l)}
	\end{equation}
In this form, the benefit of adding spoke stiffness becomes readily apparent. Because the rim stiffness and spoke stiffness act in parallel, increasing either one directly impacts the maximum tension.

@@

\begin{tabular}{c|p{12cm}}
\hline
\textit{\textbf{Term}} & \textit{\textbf{Definition}} \\
\hline
$n_s$		& Number of spokes\\
$E_s, A_s$	& Spoke Young's modulus and cross-sectional area\\
$l$			& Spoke length\\
$\alpha$	& Out-of-plane spoke angle--wall angle of a hypothetical ``cone'' of spokes\\
$\beta$		& In-plane spoke angle--measured between the in-plane projection of the spoke and
			  a radius of the wheel\\
$\n$		& Unit vector aligned with the spoke axis\\
$\nn,\nnn$	& Two unit vectors perpendicular to the spoke axis\\
$\bm{f}$	& Force exerted by a spoke on the rim\\
$\bar{\bm{f}}$& Distributed force (per-unit-length) exerted by the spokes on the rim\\
$f_1, f_2,f_3$ & Vector components of $\bm{f}$\\
$\bm{k}_f$	& Stiffness matrix for a single spoke\\
$\bar{\bm{k}}$ & Equivalent stiffness per-unit-length of the spoke system. $\bar{\bm{k}} = \bar{\bm{k}}^{el} + \bar{\bm{k}}^{tens}$\\
$\bar{\bm{k}}^{el}$ & Part of $\bar{\bm{k}}$ proportional to $E_sA_s$\\
$\bar{\bm{k}}^{tens}$ & Part of $\bar{\bm{k}}$ proportional to $T$\\
$\bar{k}_{ij}$ & Tensor components of $\bar{\bm{k}}$\\
$\e,\ee,\eee$ & Coordinate axes defined at the rim shear center\\
$s$			& Arc-length coordinate along the rim\\
$\bm{u}$	& Displacement of the rim shear center\\
$\kappa_1,\kappa_2,\kappa_3$ & Spatial curvature of the rim\\
$u,v,w$		& Vector components of $\bm{u}$ in the \e{}, \ee{}, \eee{} directions.\\
$\phi$		& Rotation of the rim cross-section about the $\hat{\bm{e}}_3$ axis.\\
$\delta u, \delta \phi$ & Virtual displacement and rotation\\
$u_n, \phi_n$ & Buckling mode coefficients\\
$n$			& Buckling mode number\\
$U_{rim}$	& Strain energy in the rim due to buckling\\
$V_{rim}$	& Virtual work of internal forces in the rim\\
$U_{spokes}$& Strain energy in the spokes due to buckling\\
$T$			& Tension in a spoke\\
$T_{cn}$	& Critical tension for the $n$th mode\\
$T_c$		& Critical buckling tension (minimum of $T_{cn}$ over all modes)\\
$T_e$		& ``Euler'' tension--tension which would produce an axial force in the rim
			  equal to the buckling load for a straight beam-column of length $2\pi R$\\
$R$			& Rim radius--measured from the rim shear center\\
$E$			& Rim Young's modulus\\
$I_{11}$	& Second moment of area of the rim in the lateral direction\\
$J$			& Rim torsion constant\\
$I_w$		& Rim warping constant\\
$\widetilde{GJ}$ & Effective torsional rigidity including both Saint-Venant torsion and warping\\
$\tilde{\mu}$ & Ratio of effective torsional rigidity to bending rigidity: $\widetilde{GJ}/EI_{11}$\\
$\lambda$	& Ratio of spoke system lateral stiffness to rim bending stiffness OR ratio
			  of rim radius to the characteristic length of a beam on an elastic foundation.\\
\hline
\end{tabular}
